---
title: "Model - Yule Simon for DD"
author: "Gabrielle Lemire"
date: "3/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(rlist)
library(reshape2)
library(rstanarm)
library(boot)
library(poweRlaw) #compare dists
library(GLDEX) #remove zeros
# load data
# load objects created in "sampling_sim.R"
# ----for Gabby
# sim_dist_100_dd = read.csv("C://Code//msu//gra//BayesNet//Paper8//sim_dists-degree_dist.csv")
# ----for Nicole
sim_dist_100_dd = read.csv("sim_dists-degree_dist.csv")

# --------------- MELT ALL PROPORTIONS
melt_all_props = function(data = sim_dist_100){
  library(reshape2)
  # create one master df with the following columns:
  # X - the size of the component
  # variable - which sim it came from
  # value - count of components

  beg_prop = seq(from = 1, to = 316, by = 21)
  end_prop = seq(from = 21, to = 336, by = 21)
  samp_prop = seq(from = .95, to = 0.2, by = -0.05)
  df = data.frame()
  df_all_sim = data.frame()
  for (i in 1:length(beg_prop)){
    beg = beg_prop[i]
    end = end_prop[i]
    df = as.data.frame(data[beg:end,])
    df$X = c(1:21) #leave as 1 to 21 because all props should have this degree count
    df = melt(df, id.vars="X")
    df_all_sim = rbind(df_all_sim, df)
  }
  return(df_all_sim)
}

# melt dataframe with all proportions and simulations into one master df_all_sim
df_all_sim_dd = melt_all_props(data = sim_dist_100_dd)

# Data for one simulation of prop=.95
data_1_dd = df_all_sim_dd[1:21,3]
data_1_dd = data_1_dd[order(-data_1_dd)]
length(data_1_dd)

# Data for one simulation of prop=.95 adding 1's
data_2_dd = df_all_sim_dd[1:21,3] + 1
data_2_dd = data_2_dd[order(-data_2_dd)]
length(data_2_dd)
```

## Overview

While discussing results from running comparisons with the poweRlaw package between four distributions (power law, log normal, exponential and poisson) we found that power law and log normal both fit different parts of the curve better and exponential and poisson have a poor fit. We want to see if adding another degree of freedom will allow the fit of the curve to be better and want to explore the option of Yule-Simon.

We have yet to explore whether this implementation (with the VGAM package) is a one or two parameter YS distribution but based on the visual for the YS on the wikipedia article (has a curve on the log-log scale) we expect it to fit differently than the power law and log normal. Whether YS or PL/LN have a better fit is yet to be determined.

We start with one simulation (with no zeros) and have so far found that for the degree distribution the model is not running (but the data formatted in the same way for the component distribution is running). *Why is that?* 

## Yule-Simon Model Implementation for one Simulation

```{r, echo=FALSE}
library(VGAM)
data_dd = data.frame(degree = 1:21, count = data_2_dd)

model_dd <- VGAM::vglm(count ~ degree, yulesimon, data = data_dd, trace = TRUE)
# https://www.rdocumentation.org/packages/VGAM/versions/0.9-0/topics/yulesimon
summary(model_dd)
coef(model, matrix = TRUE)
VGAM::plotvglm(model)
head(fitted(model))
```

## Random Code for Reference

```{r}
# random example to try from cran notes for VGAM
dyules(1:20, 2.1)
ryules(20, 2.1)

round(1000 * dyules(1:8, 2))
table(ryules(1000, 2))

## Not run:  
x <- 0:6
plot(x, dyules(x, shape = 2.2), type = "h", las = 1, col = "blue") 
## End(Not run)

# example from
# https://www.rdocumentation.org/packages/VGAM/versions/0.9-0/topics/yulesimon
ydata <- data.frame(x2 = runif(nn <- 1000))
ydata <- transform(ydata, y = ryules(nn, rho = exp(1.5 - x2)))
with(ydata, table(y))
fit <- vglm(y ~ x2, yulesimon, ydata, trace = TRUE)
coef(fit, matrix = TRUE)
summary(fit)
```